---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Elisabeth Winternitz (12641423)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Exercise 1: Initializing git (4 Points)

### a)
Completed.

### b)
Completed.

### c)
Noted. Will commit the file after completion of each subtask.

### d)
2 strengths:
- The history makes it easy to understand changes or revert them if there are problems.
- Git is great for collaboration.

2 weaknesses:
- Git has a rather steep learning curve.
- Git is not great for storing data/ with large files.

### e)
Knitted. 

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

### a)
Link to the repository: https://github.com/summer-murmuration/exeRcise-sheet-3

### b)
Code pushed (through terminal).


### c)
Noted. I will regularly push the latest changes to GitHub, especially after finished with the sheet.

## Exercise 3: Pixar Films (4 Points)

### a)

```{r}
library(tidyverse)
#load dataset into R
pixar_films <- read_csv("data/pixar_films.csv")

#clean dataset by removing films w/o a title
library(dplyr)
pixar_films <- pixar_films%>% 
  filter(!is.na(film))

#Inspect variable film_rating
unique(pixar_films$film_rating)
```
*film_rating*: the variable film_rating has, in this dataset, two values: G and PG (three films have N/A values for this variable). This refers to the Motion Picture Association film rating system. G means the film is rated for general audiences, and all ages are admitted. PG means that parental guidance is suggested, as some material may not be suitable for children.

```{r}
#Create a factor variable for the film rating:
pixar_films$film_rating_asfactor <- factor(pixar_films$film_rating,
                                         levels = c("G", "PG"),
                                         ordered = TRUE)

```
*Why is this appropriate?* 
The film rating system has a logical order, which means that "ordering" the variable (PG is more restricted than G) as a factor variable is useful for working further with this dataset. 

b)  Inspect the film titles manually. Which films form a film series? A
    film series can be identified by a common word in the titles of the
    films, often in conjunction with a number in the title,
    e.g. "Despicable Me" and "Despicable Me 2". Create a dataframe which
    displays a list of the different series with the titles of the films
    and how many films belong to the series. Output the dataframe.

c)  Load the `public_response` dataframe into R. Convert the
    `cinema_score` variable into a factor while ensuring the factor
    levels are defined in ascending order, from the lowest to the
    highest score. Combine `public_response` with the `pixar_films`
    dataset using an appropriate merge variable.

d)  Choose one of the variables representing the public response and
    create a bar plot for the films belonging to a series. Here are the
    details of the plot:

    -   The film series are represented on the x-axis.
    -   Your chosen public response variable is displayed on the y-axis.
    -   Each film in the series is represented as a separate bar. Bars
        are grouped by film under their respective series on the x-axis.
        Order the bars within a series according to the release date of
        the films.
    -   A title and axis labels for context.

    What do you notice when comparing the scores of the films in a
    series? Do you see any patterns?

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
b)  Download the dataset and track it in git.
c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: ...
-   Link: <https://github.com/owid/owid-datasets/>...

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
e)  Use R to answer your chosen question. Please limit your analysis to
    the functions and techniques we have covered so far in the course.
    You are **not expected** to use advanced statistical models or
    external packages which haven't been introduced.
f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.

## Final Note

Make sure to push all your commits and changes to GitHub before
submitting the exercise sheet.
