---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Elisabeth Winternitz (12641423)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Exercise 1: Initializing git (4 Points)

### a)
Completed.

### b)
Completed.

### c)
Noted. Will commit the file after completion of each subtask.

### d)
2 strengths:
- The history makes it easy to understand changes or revert them if there are problems.
- Git is great for collaboration.

2 weaknesses:
- Git has a rather steep learning curve.
- Git is not great for storing data/ with large files.

### e)
Knitted. 

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

### a)
Link to the repository: https://github.com/summer-murmuration/exeRcise-sheet-3

### b)
Code pushed (through terminal).


### c)
Noted. I will regularly push the latest changes to GitHub, especially after finished with the sheet.

## Exercise 3: Pixar Films (4 Points)

### a)

```{r load pixar into R}
library(tidyverse)
#load dataset into R
pixar_films <- read_csv("data/pixar_films.csv")

#clean dataset by removing films w/o a title
library(dplyr)
pixar_films <- pixar_films%>% 
  filter(!is.na(film))

#Inspect variable film_rating
unique(pixar_films$film_rating)
```
*film_rating*: the variable film_rating has, in this dataset, two values: G and PG (three films have N/A values for this variable). This refers to the Motion Picture Association film rating system. G means the film is rated for general audiences, and all ages are admitted. PG means that parental guidance is suggested, as some material may not be suitable for children.

```{r film rating variable}
#Create a factor variable for the film rating:
pixar_films$film_rating_asfactor <- factor(pixar_films$film_rating,
                                         levels = c("G", "PG"),
                                         ordered = TRUE)

```
*Why is this appropriate?* 
The film rating system has a logical order, which means that "ordering" the variable (PG is more restricted than G) as a factor variable is useful for working further with this dataset. 

### b)
*Which films form a film series?*
- Toy Story (original, 2, 3, and 4)
- Monsters, Inc and Monsters University
- Finding Nemo and Finding Dory
- Cars (original, 2, and 3)
- The Incredibles (original and 2)

```{r dataframe creation}
#Create a data frame which displays a list of the different series with the titles of the films and how many films belong to the series.

#Manually defining the film series:
list_of_series <- list(
  "Toy Story" = c("Toy Story", "Toy Story 2", "Toy Story 3", "Toy Story 4", "Lightyear"),
  "Cars" = c("Cars", "Cars 2", "Cars 3"),
  "Finding Nemo" = c("Finding Nemo", "Finding Dory"),
  "The Incredibles" = c("The Incredibles", "Incredibles 2"),
  "Monsters" = c("Monsters, Inc.", "Monsters University")
)

#Convert to a dataframe
library(dplyr)

pixar_series_df <- bind_rows(
  lapply(names(list_of_series), function(series_name) {
    films <- list_of_series[[series_name]]
    data.frame(
      series = series_name,
      film = films,
      n_films_in_series = length(films), #creating no. of films in series variable
      stringsAsFactors = FALSE
    )
  })
)

#Output the dataframe

print(pixar_series_df, n=10, na.print = "NA")
```


### c)  
```{r load public response}
#Load 'public_response' into R:
library(tidyverse)
public_response <- read_csv("data/public_response.csv")

#Convert 'cinema_score' into a factor while ensuring the factor levels are defined in ascending order from lowest to highest:
#What scores are represented in our dataset?
unique(public_response$cinema_score)
#Only 3: A+, A, and A-

score_levels <- c("A-", "A", "A+") #in order

public_response$cinema_score <- factor(public_response$cinema_score,
                                       levels = score_levels,
                                       ordered = TRUE)

#Combine 'public_response' with 'pixar_films' dataset using merge variable (film title)
combined_films_response_df <- merge(pixar_films, public_response, by = "film", all.x = TRUE)

#all.x = TRUE ensures that "Lightyear" and "Turning Red" aren't left out of the merged dataset.

```


### d)
```{r bar plot}
#Create bar plot w/ (i) film series on x-axis, (ii) chosen variable on y-axis, (iii) each film as seperate bar, (iv) w/ bars grouped by series on x-axis and (v) bars ordered by release date of films; (vi) include title and axis labels.

#load packages
library(ggplot2)
library(ggh4x)
library(dplyr)

#add series data to dataframe:
combined_films_response_df_ws <- merge(combined_films_response_df, pixar_series_df, by = "film", all.x = TRUE)

#filter for films in a series
series_df <- combined_films_response_df_ws %>%
  filter(!is.na(series)) %>% #remove films not in a series
  arrange(series, release_date) %>% #arrange by release date; (v) fulfilled
  mutate(film_order = factor(film, levels = unique(film)))

#nested factor for the x axis
series_df <- series_df %>%
  mutate(nested_x = interaction(series, film, sep = "___"))
#helps the axis understand which films belong to which series to create nested (two-level) axis labels.

#create barplot
ggplot(series_df, aes(x = film_order, y = rotten_tomatoes, fill = series)) + #film series on x axis (i), chosen variable on y axis (ii); (i) and (ii) fulfilled.
  geom_col(width = 0.7) + #each film has a seperate, equally sized bar. (iii) fulfilled.
  geom_text(aes(label = rotten_tomatoes), vjust = -0.5, size = 3) +
  scale_fill_manual(values = c(
    "Cars" = "darkred",
    "Finding Nemo" = "#e85b26",
    "Monsters" = "#9ab357",
    "The Incredibles" = "#d52d12",
    "Toy Story" = "#77c0e3"
  )) +
  scale_x_discrete(
    guide = guide_axis_nested( #I did some research on how I might best display multiple tidbits of information on one axis, and ended up finding this function within the ggplot "universe" (extension ggh4x)
      nested_levels = list(~series, ~film) #defines levels of nested axis
    ),
    labels = function(x) {
      # Match x to film names
      series_df$film[match(x, series_df$film_order)]
    } #films grouped by series; nested axis allows both title and series to "coexist". (iv) fulfilled.
  ) +
  labs(title = "Rotten Tomatoes Scores by Pixar Film Series",
       x = NULL,
       y = "Rotten Tomatoes Score",
       fill = "Series") + #labels; (vi) fulfilled.
  theme_minimal() +
  theme(
    axis.text.x.top = element_text(size = 12, face = "bold", vjust = 0.5),
    axis.text.x = element_text(size = 10, angle = 45, vjust = 1, hjust = 1),
    axis.ticks.x = element_blank(),
    legend.position = "bottom" #make sure the series are represented at the bottom, so it's easier to understand the plot
  )
```


*What do you notice when comparing the scores of the films in a series? Do you see any patterns?*
Generally, there is a trend that films after the first in the series perform worse in the Rotten Tomatoes score, meaning the score on average declines for new films in the series, in some series (Cars) more drastically than others (Toy Story).

## Exercise 4: Open Analysis (4 points)

### a)
### b)
Dataset downloaded, tracked in git.

### c)

-   Dataset Name: Estimated average age at marriage, by gender (UN and OECD)
-   Link: <https://github.com/owid/owid-datasets/tree/master/datasets/Estimated%20average%20age%20at%20marriage%2C%20by%20gender%20(UN%20and%20OECD)>

### d)
*Reasearch questions*: How has the average age at marriage developed over time, for both men and women, around the world?

This question interests me because I want to see if the "gap" in average age at marriage between men and women has started to decline, or if it remains as both men and women tend to get married later in the modern world.

### e), f)
*Note*: as my question is best answered graphically, I will work on answering it for both subtasks 4e) and 4f).

```{r data wrangling}
library(tidyverse)
library(dplyr)


#load dataset into R
marriage_data <- read_csv("data/marriage_data.csv")

#I'm interested in yearly averages over time, so I need to get a new dataframe where I have yearly averages, seperated by gender.
#First, though, I want to make sure that I only use data where a country had data for both men and women available:
marriage_cleaned <- marriage_data %>%
  filter(!is.na(`Estimated average age at marriage, women`), !is.na(`Estimated average age at marriage, men`))

#Now, to calculate yearly averages; cleaning should have removed any na values, but I am keeping the na.rm redundancy as a 'failsafe', considering the dataset is large.
yearly_avg_by_gender <- marriage_cleaned %>%
  group_by(Year) %>%
  summarise(
    avg_age_women = round(mean(`Estimated average age at marriage, women`, na.rm = TRUE), 2),
    avg_age_men = round(mean(`Estimated average age at marriage, men`, na.rm = TRUE), 2)
  )

#Okay, but I need to do a bit work so that I can create the plot I have in mind to answer my question. Namely, I want two lines by gender; I have to make the data longer.
yr_avg_long <- yearly_avg_by_gender %>%
  pivot_longer(
    cols = c(avg_age_women, avg_age_men),
    names_to = "Gender",
    values_to = "average_age"
  )
  
#To make the legend more legible, I want to rename the values of "Gender":
yr_avg_long <- yr_avg_long %>%
  mutate(Gender = recode(Gender,
                         "avg_age_men" = "Men",
                         "avg_age_women" = "Women"))

#Now I have my data properly in order, I will create a plot. I want to do this in another chunk so that I can provide a figure caption!
```

```{r createplot, fig.cap="Average age at marriage over time by gender"}
library(ggplot2)
library(bookdown)

ggplot(data = yr_avg_long, aes(
  x = Year,
  y = average_age,
  color = Gender)) +
    geom_line(linewidth = 1.2) + #allows us to see differences in gender graphically
  scale_color_manual(values = c("Men" = "darkblue", "Women" = "deeppink3")) +
    labs(
      title = "Average Age at Marriage, Over Time, by Gender",
      x = "Year",
      y = "Average Age",
      color = "Gender"
    ) +
    theme_minimal()
```

In Figure 1, the average age at marriage is displayed in two lines, separated by gender. It shows how this trend has developed over time.


