---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Elisabeth Winternitz (12641423)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Exercise 1: Initializing git (4 Points)

### a)
Completed.

### b)
Completed.

### c)
Noted. Will commit the file after completion of each subtask.

### d)
2 strengths:
- The history makes it easy to understand changes or revert them if there are problems.
- Git is great for collaboration.

2 weaknesses:
- Git has a rather steep learning curve.
- Git is not great for storing data/ with large files.

### e)
Knitted. 

## Exercise 2: Putting your Repository on GitHub (3 Points)

For this task you will upload your solution to GitHub.

### a)
Link to the repository: https://github.com/summer-murmuration/exeRcise-sheet-3

### b)
Code pushed (through terminal).


### c)
Noted. I will regularly push the latest changes to GitHub, especially after finished with the sheet.

## Exercise 3: Pixar Films (4 Points)

### a)

```{r}
library(tidyverse)
#load dataset into R
pixar_films <- read_csv("data/pixar_films.csv")

#clean dataset by removing films w/o a title
library(dplyr)
pixar_films <- pixar_films%>% 
  filter(!is.na(film))

#Inspect variable film_rating
unique(pixar_films$film_rating)
```
*film_rating*: the variable film_rating has, in this dataset, two values: G and PG (three films have N/A values for this variable). This refers to the Motion Picture Association film rating system. G means the film is rated for general audiences, and all ages are admitted. PG means that parental guidance is suggested, as some material may not be suitable for children.

```{r}
#Create a factor variable for the film rating:
pixar_films$film_rating_asfactor <- factor(pixar_films$film_rating,
                                         levels = c("G", "PG"),
                                         ordered = TRUE)

```
*Why is this appropriate?* 
The film rating system has a logical order, which means that "ordering" the variable (PG is more restricted than G) as a factor variable is useful for working further with this dataset. 

### b)
*Which films form a film series?*
- Toy Story (original, 2, 3, and 4)
- Monsters, Inc and Monsters University
- Finding Nemo and Finding Dory
- Cars (original, 2, and 3)
- The Incredibles (original and 2)

```{r}
#Create a data frame which displays a list of the different series with the titles of the films and how many films belong to the series.

#Manually defining the film series:
list_of_series <- list(
  "Toy Story" = c("Toy Story", "Toy Story 2", "Toy Story 3", "Toy Story 4", "Lightyear"),
  "Cars" = c("Cars", "Cars 2", "Cars 3"),
  "Finding Nemo" = c("Finding Nemo", "Finding Dory"),
  "The Incredibles" = c("The Incredibles", "Incredibles 2"),
  "Monsters" = c("Monsters, Inc.", "Monsters University")
)

#Convert to a dataframe
library(dplyr)

pixar_series_df <- bind_rows(
  lapply(names(list_of_series), function(series_name) {
    films <- list_of_series[[series_name]]
    data.frame(
      series = series_name,
      film = films,
      n_films_in_series = length(films),
      stringsAsFactors = FALSE
    )
  })
)

#Output the dataframe

print(pixar_series_df, n=10)
```


### c)  
```{r}
#Load 'public_response' into R:
library(tidyverse)
public_response <- read_csv("data/public_response.csv")

#Convert 'cinema_score' into a factor while ensuring the factor levels are defined in ascending order from lowest to highest:
#What scores are represented in our dataset?
unique(public_response$cinema_score)
#Only 3: A+, A, and A-

score_levels <- c("A-", "A", "A+")

public_response$cinema_score <- factor(public_response$cinema_score,
                                       levels = score_levels,
                                       ordered = TRUE)

#Combine 'public_response' with 'pixar_films' dataset using merge variable (film title)
combined_films_response_df <- merge(pixar_films, public_response, by = "film", all.x = TRUE)

#all.x = TRUE ensures that "Lightyear" and "Turning Red" aren't left out of the merged dataset.

```


### d)
```{r}
#Create bar plot w/ (i) film series on x-axis, (ii) chosen variable on y-axis, (iii) each film as seperate bar, (iv) w/ bars grouped by series on x-axis and (v) bars ordered by release date of films; (vi) include title and axis labels.

#load packages
library(ggplot2)
library(dplyr)

#add series data to dataframe:
combined_films_response_df_ws <- merge(combined_films_response_df, pixar_series_df, by = "film", all.x = TRUE)

#filter for films in a series
series_df <- combined_films_response_df_ws %>%
  filter(!is.na(series)) %>% #remove films not in a series
  arrange(series, release_date) %>% #arrange by release date; (v) fulfilled
  mutate(film_order = factor(film, levels = unique(film)))

#nested factor for the x axis
series_df <- series_df %>%
  mutate(nested_x = interaction(series, film, sep = "___"))
#helps the axis understand which films belong to which series to create nested (two-level) axis labels.

#create barplot
ggplot(series_df, aes(x = film_order, y = rotten_tomatoes, fill = series)) + #film series on x axis (i), chosen variable on y axis (ii); (i) and (ii) fulfilled.
  geom_col(width = 0.7) + #each film has a seperate, equally sized bar. (iii) fulfilled.
  geom_text(aes(label = rotten_tomatoes), vjust = -0.5, size = 3) +
  scale_fill_manual(values = c(
    "Cars" = "darkred",
    "Finding Nemo" = "#e85b26",
    "Monsters" = "#9ab357",
    "The Incredibles" = "#d52d12",
    "Toy Story" = "#77c0e3"
  )) +
  scale_x_discrete(
    guide = guide_axis_nested(
      nested_levels = list(~series, ~film) #defines levels of nested axis
    ),
    labels = function(x) {
      # Match x to film names
      series_df$film[match(x, series_df$film_order)]
    } #films grouped by series; nested axis allows both title and series to "coexist". (iv) fulfilled.
  ) +
  labs(title = "Rotten Tomatoes Scores by Pixar Film Series",
       x = NULL,
       y = "Rotten Tomatoes Score",
       fill = "Series") + #labels; (vi) fulfilled.
  theme_minimal() +
  theme(
    axis.text.x.top = element_text(size = 12, face = "bold", vjust = 0.5),
    axis.text.x = element_text(size = 10, angle = 45, vjust = 1, hjust = 1),
    axis.ticks.x = element_blank(),
    legend.position = "bottom" #make sure the series are represented at the bottom, so it's easier to understand the plot
  )
```


*What do you notice when comparing the scores of the films in a series? Do you see any patterns?*
Generally, there is a trend that films after the first in the series perform worse in the Rotten Tomatoes score, meaning the score on average declines for new films in the series, in some series (Cars) more drastically than others (Toy Story).

## Exercise 4: Open Analysis (4 points)

This exercise is a bit more open-ended. You can choose any dataset from
[Our World in Data](https://ourworldindata.org/) and analyze it, while
determining the research question yourself.

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets>
    and choose a dataset that interests you. You can have a look at
    <https://ourworldindata.org/> to gather some inspiration.
b)  Download the dataset and track it in git.
c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: ...
-   Link: <https://github.com/owid/owid-datasets/>...

d)  Come up with a (research) question you want to answer with the data
    and briefly explain why you believe this is an interesting question
    within one sentence. It should be a question that can be answered
    with the dataset and using R.
e)  Use R to answer your chosen question. Please limit your analysis to
    the functions and techniques we have covered so far in the course.
    You are **not expected** to use advanced statistical models or
    external packages which haven't been introduced.
f)  Create a meaningful plot / figure with the dataset. Make sure to
    provide a figure caption (via the chunk options / Rmarkdown) and
    correctly label the figure.

## Final Note

Make sure to push all your commits and changes to GitHub before
submitting the exercise sheet.
